---
title: "Insurance Expenses Analysis"
output: html_document
date: "2022-12-26"
---

Copy the dataset to your working directory then load the data .

```{r}
insurance <- read.csv("insurance.csv" , stringsAsFactors = TRUE)
head(insurance)

```

Lets have a look at the structure of the data
```{r}
str(insurance)
```

 We need to change the name of the target variable to expenses.

```{r}
library(dplyr)
insurance = insurance %>%
  rename(expenses = charges)
```

It is important to check for normality.
```{r}
summary(insurance$expenses)
```
 NB: the expenses are skewed to the right because the mean > the median.

Plotting the results
```{r}
hist(insurance$expenses)
```

How are the policyholders distributed across the regions?
```{r}
table(insurance$region)

```

What are the relationships in the among the feature variables and the dependent variable?
```{r}
cor(insurance[c("age", "bmi", "children", "expenses")])
```


Let's produce a scatterplot matrix for these relationships
```{r}
pairs(insurance[c("age","bmi","children","expenses")])
```


Lets add more info to the plots
```{r}
library(psych)
pairs.panels(insurance[c("age", "bmi","children","expenses")])
```


The next step is to build a model
```{r}
ins_model <-  lm(expenses ~ age + sex + bmi + children + smoker + region ,data = insurance)
ins_model
```


How well does our model fit the data?
```{r}
summary(ins_model)
```
The adjusted R-squared value is 0.74 . 

The model is performing quite well but can the perfornmace be improved ?

 .age has a non-linear relationship on medical expenses
```{r}
insurance$age2 <- insurance$age^2
```


.The bmi's effect on expenses is significant after the 30 threshold i.e only if individual is obese
```{r}
insurance$bmi30 <- ifelse(insurance$bmi >= 30 , 1 , 0)
```


.We will also add the interaction (combined effect) of smoker and obesity variables 

hence our improved model is
```{r}
ins_model2 <- lm(expenses ~ age + age2 + sex + bmi + children + region + smoker*bmi30 ,data = insurance)
summary(ins_model2)
```
The model performs better than the previous model 
The adjusted R-squared has increased from 0.75 to 0.87

Now we can perform some predictions on the training data
```{r}
insurance$pred <-  predict(ins_model2 , insurance)
```


What is the correlation between the predicted and actual cost of insurance
```{r}
cor(insurance$pred , insurance$expenses) 
```
0.93 indicates a very strong linear relationship between predicted and actual values.

lets display this graphically . 
```{r}
plot(insurance$pred , insurance$expenses)
abline(a = 0 , b = 1 , col = "red", lwd = 3 , lty = 2)
```


In the real world we could have a propective policyholder with the following characteristics:
 
 30 year old, overweight, male non-smoker with two children in the Northeast.
 
 The predicted expense is :
```{r}
predict(ins_model2 , data.frame(age = 30 , age2 = 30^2 , bmi = 30 ,bmi30 = 1 , sex = "male" , 
                                smoker = "no" , children = 2 ,region = "northeast"))

```

The model expects the company to pay 5972.859 in expenses . 

Comparing this rate to a female with the same characteristics as the male counterpart : 

```{r}
predict(ins_model2 , data.frame(age = 30 , age2 = 30^2 , bmi = 30 , bmi30 = 1 , sex = "female" ,
                                 smoker = "no" , children = 2 , region = "northeast"))
```

The model expects the company to pay 6469.683 in expenses .

What is the effect of reduction in  thee number of children from two to zero on expenses ?
```{r}
predict(ins_model2 , data.frame(age = 30 , age2 = 30^2 , bmi = 30 , bmi30 = 1 , sex = "female" , 
                                smoker = "no" , children = 0 ,region = "northeast"))
```

the expenses reduce to 5112.561   .

Following similar steps for a number of additional customer segments, the insurance
company would be able to develop a profitable pricing structure for various demographics